---
title: "Data Science II Homework 6"
author: "Charlotte Abrams"
date: "5/07/2019"
output: github_document
---
___


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
library(factoextra)
library(gridExtra)
library(corrplot)
library(RColorBrewer)
library(gplots)
library(ape)
library(RColorBrewer)
library(dplyr)
library(purrr)
```

## Cluster Analysis
We perform hierarchical clustering on the states using the USArrests data in the ISLR
package. For each of the 50 states in the United States, the data set contains the number
of arrests per 100,000 residents for each of three crimes: Assault, Murder, and Rape. The
data set also contains the percent of the population in each state living in urban areas,
UrbanPop. The four variables will be used as features for clustering.

```{r}
arrests = USArrests
```

### Question 1 (a)
Using hierarchical clustering with complete linkage and Euclidean distance, cluster the
states.

```{r}
#Summarize data
summary(arrests)

set.seed(1)

hc.complete <- hclust(dist(arrests), method = "complete")

fviz_dend(hc.complete, k = 4,
          cex = 0.3,
          palette = "jco",
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, rect_border = "jco",
          labels_track_height = 2.5)

```

### Question 1(b)
Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

```{r}
set.seed(1)
hc.complete.cut <- cutree(hc.complete, 3)
#cluster 1
arrests[hc.complete.cut == 1,]
#cluster 2
arrests[hc.complete.cut == 2,]
#cluster 3
arrests[hc.complete.cut == 3,]
```

### Question 1(c)
Hierarchically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one.
```{r}
set.seed(1)
sd.data <- scale(arrests)

hc.complete.sd <- hclust(dist(sd.data), method = "complete")

fviz_dend(hc.complete.sd, k = 4,
          cex = 0.3,
          palette = "jco",
          color_labels_by_k = TRUE,
          rect = TRUE, rect_fill = TRUE, rect_border = "jco",
          labels_track_height = 2.5)
```

### Question 1(d)
What effect does scaling the variables have on the hierarchical clustering obtained? In
your opinion, should the variables be scaled before the inter-observation dissimilarities are
computed?

-  Scaling the variables affects the clusters created. We can see that the clusters in Part (a) and Part (c) are different due to scaling the variable. I think it would be better to scale the variables before the inter-observation dissimilarities are computed. This is because different variables have different units, so it's best to standardize everything earlier so we can compare them on equal grounds.

## PCA
PCA can be used for image compression. In this question, we use the jpeg package to
read and write the .jpeg files. We use a image of cat for illustration, and the sample codes
are given in "image.R". Read the image using img <- readJPEG(`example.jpg'). The
image will be represented as three matrices as an array with each matrix corresponding
to the RGB color value scheme and each element in a matrix corresponding to one pixel.
Extract the individual color value matrices to perform PCA on each of them. Reconstruct
the original image using the projections of the data with the first 20 PCs.
Now use your own .jpg image to perform image compression via PCA with different numbers
of PCs (e.g., 50, 100, 200, ...).

```{r}
library(jpeg)

img <- readJPEG('corgi.jpg')

dim(img)

r <- img[,,1]
g <- img[,,2]
b <- img[,,3]

img.r.pca <- prcomp(r, center = FALSE)
img.g.pca <- prcomp(g, center = FALSE)
img.b.pca <- prcomp(b, center = FALSE)

rgb.pca <- list(img.r.pca, img.g.pca, img.b.pca)

# Approximate X with XV_kV_k^T
compress <- function(pr, k)
{
  compressed.img <- pr$x[,1:k] %*% t(pr$rotation[,1:k])
  compressed.img
}

# Using first 20 PCs
pca20 <- sapply(rgb.pca, compress, k = 20, simplify = "array")

writeJPEG(pca20, "pca20.jpeg")

# Using first 10 PCs
pca10 <- sapply(rgb.pca, compress, k = 10, simplify = "array")

writeJPEG(pca10, "pca10.jpeg")

# Using first 300 PCs
pca300 <- sapply(rgb.pca, compress, k = 300, simplify = "array")

writeJPEG(pca300, "pca300.jpeg")
```

