___
title: "P8106_hw2"
author: "Charlotte Abrams"
date: "3/15/2019"
output: github_document
___

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(boot)
library(glmnet)
library(pls)
library(splines)
library(mgcv)
```
#Import Data
```{r}
concrete <- read.csv("~/Documents/Charlotte/Columbia/Spring 2019/DSII/HW2/Data/concrete.csv")
```

# Matrix of predictors
```{r}
x <- model.matrix(CompressiveStrength~.,concrete)[,-1]
```

#Vector of response
```{r}
y <- concrete$CompressiveStrength
```
#  A  ###############################################################################################
```{r, fig.height=5}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
```

#Create scatter plots of response vs. predictors using the function featurePlot()
```{r}
featurePlot(x, y, plot = "scatter", labels = c("","Compressive Strength"),
            type = c("p"), layout = c(4, 2))
```


#  B  ###############################################################################################
#Perform polynomial regression to predict compressive strength using water as the
#predictor. For 1 < d < 4, use cross-validation to select the optimal degree d for
#the polynomial. What degree was chosen, and how does this compare to the results
#of hypothesis testing using ANOVA? Make a plot of different polynomial fits to the
#data.
```{r}
fit1 <- lm(CompressiveStrength~Water, data = concrete) #First degree
fit2 <- lm(CompressiveStrength~poly(Water,2), data = concrete) #Squared
fit3 <- lm(CompressiveStrength~poly(Water,3), data = concrete) #Cubed
fit4 <- lm(CompressiveStrength~poly(Water,4), data = concrete) #Fourth power

plot1 <- ggplot(data = concrete, aes(x = Water, y = CompressiveStrength)) +
  geom_point(color = rgb(.2, .4, .2, .5))

plot1

plot(fit1) + plot(fit2) + plot(fit3) + plot(fit4) 
```

#Calculate the 10-fold CV MSE using the functiontrain().
# 10-fold CV
```{r}
ctrl <- trainControl(method = "cv", number = 10)

set.seed(1)
lmFit1 <- train(CompressiveStrength~Water,
               data = concrete,
               method = "lm",
               trControl = ctrl)
lmFit1
```
# RMSE      Rsquared    MAE     
# 15.99058  0.09107776  13.08764
```{r}
set.seed(1)
lmFit2 <- train(CompressiveStrength~poly(Water,2),
                data = concrete,
                method = "lm",
                trControl = ctrl)
lmFit2
```
# RMSE      Rsquared   MAE     
# 15.52028  0.1460514  12.46876
```{r}
set.seed(1)
lmFit3 <- train(CompressiveStrength~poly(Water,3),
                data = concrete,
                method = "lm",
                trControl = ctrl)
lmFit3
```
# RMSE      Rsquared   MAE     
# 15.17431  0.1801495  12.25177
```{r}
set.seed(1)
lmFit4 <- train(CompressiveStrength~poly(Water,4),
                data = concrete,
                method = "lm",
                trControl = ctrl)
lmFit4
```
# RMSE      Rsquared  MAE     
# 15.02604  0.199201  12.09692
```{r}
anova(fit1,fit2,fit3,fit4)
```
#Res.  Df    RSS     Df   Sum of Sq    F         Pr(>F)    
#1     1028  263085                                  
#2     1027  247712  1    15372.8      68.140    4.652e-16 ***
#3     1026  235538  1    12174.0      53.962    4.166e-13 ***
#4     1025  231246  1    4291.5       19.022    1.423e-05 ***

#The Mean Square Error (RMSE) is the smallest when ussing the Y~Water^4 model (RMSE=15.02604), 
#therefore, it is the best model to use for this analysis. The models with the highest RMSE values and 
#the worst models for this adanlysis would be the Y~Water model with RMSE=15.99058. This aligns with 
#hypothesis testing using ANOVA.


#  C  ###############################################################################################
#Fit a smoothing spline using water as the predictor for a range of degrees of
#freedom, as well as the degree of freedom obtained by generalized cross-validation,
#and plot the resulting fits. Describe the results obtained.

#Generate test data
```{r}
Waterlims <- range(concrete$Water)
Water.grid <- seq(from = Waterlims[1],to = Waterlims[2])
```

#Cross-Validation Fit
```{r}
fit.ss <- smooth.spline(concrete$Water, concrete$CompressiveStrength)
fit.ss
```
## [1] 68.88205

```{r}
pred.ss <- predict(fit.ss,
                   x = Water.grid)
pred.ss.df <- data.frame(pred = pred.ss$y,
                         Water = Water.grid)
```
#Fit ss1 (df=50)
```{r}
fit.ss1 <- smooth.spline(concrete$Water, concrete$CompressiveStrength, df = 50)
fit.ss1
```
## [1] 50.00509

```{r}
pred.ss1 <- predict(fit.ss1, 
                    x = Water.grid)
pred.ss.df1 <- data.frame(pred = pred.ss1$y, 
                          Water = Water.grid)
```
#Fit ss2 (df=60)
```{r}
fit.ss2 <- smooth.spline(concrete$Water, concrete$CompressiveStrength, df = 60)
fit.ss2
```
## [1] 59.99469

```{r}
pred.ss2 <- predict(fit.ss2, 
                    x = Water.grid)
pred.ss.df2 <- data.frame(pred = pred.ss2$y, 
                          Water = Water.grid)
```
#Fit ss3 (df=70)
```{r}
fit.ss3 <- smooth.spline(concrete$Water, concrete$CompressiveStrength, df = 70)
fit.ss3
```
## [1] 70.00499

```{r}
pred.ss3 <- predict(fit.ss3, 
                    x = Water.grid)
pred.ss.df3 <- data.frame(pred = pred.ss3$y, 
                          Water = Water.grid)
```

#Plots
```{r}
smooth_spline_fits <- plot1 + geom_line(aes(x = Water, y = pred), data = pred.ss.df1,
                                 color = rgb(.8, .1, .1, 1)) + 
  geom_line(aes(x = Water, y = pred), data = pred.ss.df2,
            color = rgb(0, 0, 1, 1)) + 
  geom_line(aes(x = Water, y = pred), data = pred.ss.df3,
            color = rgb(1, 0, 1, 1)) + 
  geom_line(aes(x = Water, y = pred), data = pred.ss.df,
            color = rgb(.2, .2, .4, 1)) + 
  theme_bw()
smooth_spline_fits
```

#Through generalized cross-validation, we found that the degree of freedom obtained
#is 68.88205. Thus, I tested out other models with dfs around 68.88205 (50, 60, and 70).
#The results are shown in the comments under each model.

#  D  ###############################################################################################
#Fit a GAM using all the predictors. Plot the results and explain your findings.
```{r}
gam.m1 <- gam(CompressiveStrength~Cement+Water+BlastFurnaceSlag+FlyAsh+Superplasticizer+CoarseAggregate+FineAggregate+Age, data = concrete)
gam.m2 <- gam(CompressiveStrength~Cement+s(Water)+BlastFurnaceSlag+FlyAsh+Superplasticizer+CoarseAggregate+FineAggregate+Age, data = concrete)

anova(gam.m1, gam.m2, test = "F")
```
#Resid. Df Resid. Dev     Df Deviance      F   Pr(>F)
#1    1021.0     110413                                
#2    1013.4     106140 7.5562   4272.8 5.4038 2.01e-06

```{r}
plot(gam.m2)

vis.gam(gam.m1, view = c("Water","FlyAsh"),
       plot.type = "contour", color = "topo")

vis.gam(gam.m2, view = c("Water","FlyAsh"),
        plot.type = "contour", color = "topo")
```
